Hydrodynamics and rheology of an adhesive vesicle doublet

### 2019-02-08 vesicle_code/docs/adhesion/adhesion.tex
first edits for Feb 08 from yny

\title{Hydrodynamics and rheology of adhesive vesicle doublets}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\begin{abstract}
The dynamics of adhesive vesicles under various flowing conditions is
investigated numerically using a high-order adaptive-in-time boundary
integral equation formulation. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a doublet at equilibrium. A lubrication
analysis on such draining of the thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet are observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the hydrodynamic
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form as they approach each other provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of both reduced area and
the adhesion strength.  Results from these numerical studies and
analysis shed light on the hydrodynamic effects of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to place the stagnation point at a desirable location by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.  To explore
the application of such a fluid trap to measure the adhesion strength
between two bound vesicle membranes, we propose the following thought
experiment.

\begin{figure}[htp]
  \begin{minipage}{0.53\textwidth}
    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
  \end{minipage}
   \caption{\label{fig:InclinationAngle} The
  definition of the inclination angle.  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the diverging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the diverging direction.  The square
  mark is a stagnation point and the round point is located at the
  center of the vesicle.}
\end{figure}

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet, which has a flat contact region as shown in \S~\ref{sec:qflow}. 
Based on how the contact region aligns with the direction of divergent flow, 
we can define the inclination angle of the vesicle doublet as illustrated in Figure~\ref{fig:InclinationAngle}.
%

When $\theta=\pi$ (the long axis of the vesicle doublet is aligned with the far-field diverging flow), the converging flow
pushes vesicles towards each other and the short-range repulsive force is essential to keep the two vesicles at the separation distance.
With the fluid flow pushing the two vesicles towards the stagnation point dynamically placed at the doublet center, 
this configuration is expected to be more stable than the configuration in the middle of Figure~\ref{fig:InclinationAngle} where $\theta=\pi/2$ 
(the long axis of vesicle doublet is parallel to the converging direction).  As shown in the illustration, 
the diverging flow is pulling the vesicles away from each other when $\theta=\pi/2$. 
Thus it is possible that the diverging flow may overcome the adhesion and
break up the doublet. It is also reasonable to expect the vesicle doublet to rotate around the stagnation so the flow is pushing them together rather than pulling them apart.

In the following simulation experiments we place the vesicle doublet with $\theta=\pi/2$ so that the flat contact region is
orthogonal to the diverging flow that is pulling vesicles away from each other.
%In this configuration the flow is pulling the vesicles away from each other, and 
%is expected to be less stable than the $\theta=\pi$ configuration where the fluid flow is constantly pushing vesicles towards each other. 
%
At low extensional
rates, we expect the vesicle doublet to stay at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains bound for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is split with an extensional rate $\chi = 0.1$
(Figure~\ref{fig:extensional4}).  
%
For extensional rates and reduced areas that result in a bound vesicle doublet, 
we observe that the doublet always rotate from the $\theta=\pi/2$ configuration towards
the more stable $\theta=\pi$ configuration, as shown in Figures~\ref{fig:extensional1}--\ref{fig:extensional3}.
We also observe that in most cases with moderate extensional flow rates the doublet falls short of 
aligning their long axis with the diverging direction.

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.02$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
  \end{figure}

>>>>

\title{Hydrodynamics and rheology of an adhesive vesicle doublet}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\begin{abstract}
The dynamics of an adhesive vesicle doublet under various flowing conditions is
investigated numerically using a high-order adaptive-in-time boundary
integral equation formulation. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a vesicle doublet at equilibrium. A lubrication
analysis on such draining of a thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is dynamically
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the 
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form as two vesicles approach each other provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of both reduced area and
the adhesion strength.  Results from these numerical studies and
analysis shed light on the hydrodynamic effects of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity at the origin) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in the orthogonal direction.  With an active control algorithm
to place the stagnation point at a desirable location by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.  To explore
the application of such a fluid trap to measure the adhesion strength
between two bound vesicle membranes, we propose the following thought
experiment.

\begin{figure}[htp]
  \begin{minipage}{0.53\textwidth}
    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
  \end{minipage}
   \caption{\label{fig:InclinationAngle} The
  definition of the inclination angle of a vesicle doublet at the center of an extensional flow.  
  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the diverging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the diverging direction.  The square
  mark is a stagnation point and the round point is located at the
  center of the vesicle.}
\end{figure}

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet, which has a flat contact region as shown in \S~\ref{sec:qflow}. 
Based on how the contact region aligns with the extensional flow, 
we can define the inclination angle of the vesicle doublet as illustrated in Figure~\ref{fig:InclinationAngle}.
%

When $\theta=\pi$ (the doublet long axis is aligned with the far-field diverging flow), the converging flow
pushes vesicles towards each other and the short-range repulsive force is essential to keep the two vesicles at the separation distance.
With the converging flow pushing the two vesicles towards the stagnation point dynamically placed at the doublet center, 
this configuration is expected to be more stable than the $\theta=\pi/2$ configuration in the middle of Figure~\ref{fig:InclinationAngle},  where  
the doublet long axis is parallel to the converging direction. This is because
when $\theta=\pi/2$ the diverging flow is pulling the vesicles away from each other as shown in the illustration. 
With the inclination angle fixed at $\theta=\pi/2$,
the vesicle doublet stays intact only when the adhesive force is strong enough to dominate over the fluid drag in this configuration.
Thus it is possible that the diverging flow may overcome adhesion and
break up the doublet. It is also reasonable to expect a vesicle doublet to rotate from $\theta=\pi/2$ to $\theta=\pi$,
such that the fluid flow is pushing them together rather than pulling them apart.

In the following numerical experiments we place a vesicle doublet with $\theta=\pi/2$, so that
the diverging flow may be strong enough to pull vesicles away from the doublet.
%In this configuration the flow is pulling the vesicles away from each other, and 
%is expected to be less stable than the $\theta=\pi$ configuration where the fluid flow is constantly pushing vesicles towards each other. 
%
At low extensional
rates, we expect the $\theta=\pi/2$ doublet to stay intact at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains bound for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is split with a critical extensional rate $\chi_c \sim 0.1$
(Figure~\ref{fig:extensional4}).  
%
\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.02$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
  \end{figure}
  %
  For extensional rates and reduced areas that result in a bound vesicle doublet, 
we observe that the doublet always rotate from $\theta=\pi/2$ configuration towards
the more stable $\theta=\pi$ configuration, as shown in Figures~\ref{fig:extensional1}--\ref{fig:extensional3}.
We also observe that in most cases with moderate extensional flow rates the doublet falls short of 
aligning their long axis with the diverging direction.
